### 2016-07-05 introduction.tex
tightened weak consistency discussion in intro

Recent results on reasoning about application invariants in the
presence of weak consistency~\cite{burckhardt14, redblueosdi,
redblueatc, ecinec, gotsmanpopl16} address broadly related concerns.
Weak consistency is a phenomenon that usually manifests on replicated
data stores, where operations are concurrently executed against
different replicas, resulting in an order of execution inconsistent
with their serial order. The operations, nonetheless, are atomic and
fully isolated, and each operation is required to locally preserve
application invariants. In contrast, weak isolation is a property of
transactions, which are sets of atomic operations. Weak isolation
manifests when successive atomic operations in a transaction witness
different contemporary states of the database (or different replicas
of the replicated store) which, although consistent individually, may
not be obviously reconciled into a consistent global view.
Intuitively, the latter problem reduces to the former if all
transactions contain a single operation. Furthermore, weak isolation
can exist independent of weak consistency, as evident from the
presence of weakly-isolated transactions on conventional RDBS. With an
added complexity of replication, richer mechanisms are needed to
reason about weak isolation in tandem with weak consistency. Thus,
frameworks for reasoning about weak isolation will necessarily have to
generalize the reasoning frameworks developed for weak consistency in
new and important ways.

>>>>

%%%SJ: tried to tighten this paragraph; the original now commented seemed overly
%%complicated
Recent results on reasoning about application invariants in the
presence of weak consistency~\cite{burckhardt14, redblueosdi,
  redblueatc, ecinec, gotsmanpopl16} address broadly related concerns.
Weak consistency is a phenomenon that manifests on replicated data
stores, where operations are concurrently executed against different
replicas, resulting in an execution order inconsistent with any serial
(sequentially consistent) ordering.  In contrast, weak isolation is a
property of atomic transactions whose effects interfere with one
another in ways inconsistent with serializability.  Unlike weak
consistency, weak isolation can manifest even in an unreplicated
setting, as evident from the support for weakly-isolated transactions
on conventional (unreplicated) databases as mentioned above.  With the
added complexity of replication, richer mechanisms are needed to
reason about weak isolation in tandem with weak consistency. Thus,
frameworks for reasoning about weak isolation will necessarily have to
generalize the reasoning frameworks developed for weak consistency in
new and important ways.

%% Recent results on reasoning about application invariants in the
%% presence of weak consistency~\cite{burckhardt14, redblueosdi,
%% redblueatc, ecinec, gotsmanpopl16} address broadly related concerns.
%% Weak consistency is a phenomenon that usually manifests on replicated
%% data stores, where operations are concurrently executed against
%% different replicas, resulting in an order of execution inconsistent
%% with their serial order. The operations, nonetheless, are atomic and
%% fully isolated, and each operation is required to locally preserve
%% application invariants. In contrast, weak isolation is a property of
%% transactions, which are sets of atomic operations. Weak isolation
%% manifests when successive atomic operations in a transaction witness
%% different contemporary states of the database (or different replicas
%% of the replicated store) which, although consistent individually, may
%% not be obviously reconciled into a consistent global view.
%% Intuitively, the latter problem reduces to the former if all
%% transactions contain a single operation. Furthermore, weak isolation
%% can exist independent of weak consistency, as evident from the
%% presence of weakly-isolated transactions on conventional RDBS. With an
%% added complexity of replication, richer mechanisms are needed to
%% reason about weak isolation in tandem with weak consistency. Thus,
%% frameworks for reasoning about weak isolation will necessarily have to
%% generalize the reasoning frameworks developed for weak consistency in
%% new and important ways.

### 2016-07-05 store-consistency.tex
tightened weak consistency discussion in intro

First, we focus on the sequential aspect of the proof and show that
the assertions that decorate \C{Wd1} are indeed valid. We consider the
triple associated with the assignment to \C{B} for illustration. The
pre-condition $\phi_4$ asserts different values for \C{B} and \C{v3}
(temporary alias of RHS expression) depending on whether or not
\C{Wd2} is committed. Inside \C{Wd1}, $\committed(\C{Wd2})$ may mean
that either \C{Wd2} happened before \C{Wd1} (hence can be visible to
all of \C{Wd1}), or that it committed concurrently with \C{Wd1} (hence
cannot be visible to all of \C{Wd2}).  Nonetheless, $\phi_4$ only
considers the case when $\C{Wd2} \visar \C{Wd1}$. As we show below,
this is sufficient to prove the post-condition $\phi_5$, which appears
to make a stronger claim about the value of \C{B}. The proof follows
from the RG rule \rulelabel{RG-Asgn} for assignments, which allows us
to easily conclude the following about the execution state after the
assignment:\vspace*{-12pt}

@@

The second part of the proof is to show that assertions are stable
despite the interference from the the concurrent thread executing
\C{Wd2}. We formalize the interference as a rely relation that
describes the semantics of \C{Wd2} at just the right level of
abstraction. In particular, we would like to capture two important
facts about \C{Wd2}: 
\begin{itemize}
  \item  If \C{Wd2} commits, then it should have already written to
  \C{B}. This fact is needed to show that \C{Wd2} conflicts with
  \C{Wd1}, so that \C{SI(Wd1)} property can be exploited to deduce
  $\C{Wd2} \visar \C{Wd1}$ from $\committed(\C{Wd2})$ in the
  sequential proof.

>>>>

First, we focus on the sequential aspect of the proof and show that
the assertions that decorate \C{Wd1} are indeed valid. We consider the
triple associated with the assignment to \C{B} for illustration. The
pre-condition $\phi_4$ asserts different values for \C{B} and \C{v3}
depending on whether or not \C{Wd2} is committed. Inside \C{Wd1},
$\committed(\C{Wd2})$ may mean that either \C{Wd2} happened before
\C{Wd1} (hence can be visible to all of \C{Wd1}), or that it committed
concurrently with \C{Wd1} (hence cannot be visible to all of \C{Wd2}).
Nonetheless, $\phi_4$ only considers the case when $\C{Wd2} \visar
\C{Wd1}$. As we show below, this is sufficient to prove the
post-condition $\phi_5$, which appears to make a stronger claim about
the value of \C{B}. The proof follows from the RG rule
\rulelabel{RG-Asgn} for assignments, which allows us to easily
conclude the following about the execution state after the
assignment:\vspace*{-12pt}

@@

The second part of the proof is to show that assertions are stable
despite the interference from the concurrent thread executing
\C{Wd2}. We formalize such interference as a rely relation that
describes the semantics of \C{Wd2} at the right level of
abstraction. In particular, we would like to capture two important
facts about \C{Wd2}: 
\begin{itemize}
  \item  If \C{Wd2} commits, then it should have already written to
  \C{B}. This fact is needed to show that \C{Wd2} conflicts with
  \C{Wd1}, so that \C{SI(Wd1)} property can be exploited to deduce
  $\C{Wd2} \visar \C{Wd1}$ from $\committed(\C{Wd2})$ in the
  sequential proof.
